Hypoglycemia Analysis - CFRD REMOVED

### 2023-03-31 Christine Chan/Hypoglycemia/tim_analysis.Rmd
Update tim_analysis.Rmd

---
title: "Hypoglycemia Analysis"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    fig_height: 8
  pdf_document:
    toc: yes
---

@@

```{r}
getAUCs <- function(data,subject,time,vars){
  all_return <- NULL

  for (var in vars){
    subs <- unique(data[[subject]])
    return <- NULL
    for (sub in subs){
      try({
        df <- data[data[[subject]] == sub,]
        ogtt_0 <- df[df[[time]] == 0,var]
        ogtt_30 <- df[df[[time]] == 30,var]
        ogtt_60 <- df[df[[time]] == 60,var]
        ogtt_90 <- df[df[[time]] == 90,var]
        ogtt_120 <- df[df[[time]] == 120,var]
        ogtt_auc <- (ogtt_30[[var]] + ogtt_60[[var]] + ogtt_90[[var]])/2 + (ogtt_0[[var]] + ogtt_120[[var]])/4
        if(length(ogtt_auc) > 0){
          row <- c(sub,ogtt_auc)
          return <- rbind(return, row)
        }

      })

    }
    return <- as.data.frame(return)
    names(return) <- c(subject,paste0(var,"_auc"))
    if (is.null(all_return)){
      all_return <- return
    }
    else{
      all_return <- merge(all_return, return, all = T)
    }

  }
  all_return

}

```

```{r, warning = F, message = F}
dat <- read.csv("./data/CGMInHealthyControls_DATA_2021-05-25_0918.csv", na = "-999")

dat$group_category[dat$group_category == "1"] <- "Healthy"
dat$group_category[dat$group_category == "2"] <- "CF-NGT"
dat$group_category[dat$group_category == "3"] <- "CF-AG"
dat$group_category[dat$group_category == "4"] <- "CFRD"
dat$group_category <- factor(dat$group_category, levels = c("Healthy","CF-NGT","CF-AG","CFRD"))

@@

dat <- dat[dat$date_of_visit != "",]

@@

gluc <- dat %>% select(subject_id, 
                       date_of_visit, 
                       age, 
                       gender, 
                       race_ethnicity, 
                       starts_with("min"),
                       -starts_with("min_spent"),
                       -contains("sensor"),
                       -contains("active"), 
                       -contains("glp"),
                       -contains("gip"))
gluc <- merge(gluc, dat %>% select(subject_id, date_of_visit, age, starts_with("ucd_")))
gluc <- unique(gluc)
gluc <- gluc[!is.na(gluc$min_30_glucose),]
gluc[] <- lapply(gluc, gsub, pattern="<", replacement = "")

gluc_names <- names(gluc)
gluc_names <- gsub("min_","",gluc_names)
gluc_names <- gsub("minute_","",gluc_names)
gluc_names <- gsub("minus_","-",gluc_names)
gluc_names <- gsub("ucd_","",gluc_names)
gluc_names <- gsub("_total","",gluc_names)
gluc_names <- gsub("glp_1","_glp1",gluc_names)
gluc_names <- gsub("c_peptide","cpeptide",gluc_names)
names(gluc) <- gluc_names
gluc[gluc == -999.99] <- NA
gluc <- unique(gluc)
new_gluc <- gluc[0,]

for (sub in unique(gluc$subject_id)){
  df <- gluc[gluc$subject_id == sub,]
  if (nrow(df) > 1){
    df <- df %>% arrange(`0_gip`,desc(date_of_visit))
    df <- df[1,]
  }
  new_gluc <- rbind(new_gluc,df)
}

gluc_wide <- new_gluc %>% select(-date_of_visit)
gluc <- gluc_wide %>% pivot_longer(cols = -c(subject_id), 
                                   names_to = c("time", ".value"), names_sep = "_")

#gluc <- gluc[!is.na(gluc$glucose) | !is.na(gluc$glucagon) | !is.na(gluc$insulin)| !is.na(gluc$cpeptide),]
#gluc <- gluc[!is.na(gluc$glucagon),]
gluc <- gluc[gluc$time != "-10",]
gluc$time <- as.factor(as.numeric(gluc$time))

@@

aucs <- getAUCs(gluc, "subject_id","time",c("glucose","glucagon","insulin","cpeptide","gip","glp1"))

df_aucs <- merge(aucs,unique(gluc %>% select(subject_id)))
df_aucs$glucagon_auc <- as.numeric(df_aucs$glucagon_auc)
df_aucs$glucose_auc <- as.numeric(df_aucs$glucose_auc)
df_aucs$insulin_auc <- as.numeric(df_aucs$insulin_auc)
df_aucs$cpeptide_auc <- as.numeric(df_aucs$cpeptide_auc)
df_aucs$gip_auc <- as.numeric(df_aucs$gip_auc)
df_aucs$glp1_auc <- as.numeric(df_aucs$glp1_auc)

df_aucs <- df_aucs[!is.na(df_aucs$glucagon_auc) | !is.na(df_aucs$glucose_auc) |!is.na(df_aucs$insulin_auc) |!is.na(df_aucs$cpeptide_auc),]

@@

gluc$hypo_60 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])),"Yes","No")
gluc$hypo_70 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])),"Yes","No")

gluc_aucs <- merge(gluc_aucs,unique(gluc %>% select(subject_id, hypo_60,hypo_70)))

glucose <- gluc_wide %>% select(subject_id, contains("glucose"),-starts_with("-10"))
glucose <- glucose[complete.cases(glucose),]
glucose[,2:10] <- sapply(glucose[,2:10], as.numeric)
glucose$glucose_AUC <- rowSums(glucose[,2:10])
glucose$glucose_mean <- glucose$glucose_AUC/9

insulin <- gluc_wide %>% select(subject_id, contains("insulin"),-starts_with("-10"))
insulin <- insulin[complete.cases(insulin),]
insulin[,2:10] <- sapply(insulin[,2:10], as.numeric)
insulin$insulin_AUC <- rowSums(insulin[,2:10])
insulin$insulin_mean <- insulin$insulin_AUC/9

cpeptide <- gluc_wide %>% select(subject_id, contains("cpeptide"),-starts_with("-10"))
cpeptide <- cpeptide[complete.cases(cpeptide),]
cpeptide[,2:10] <- sapply(cpeptide[,2:10], as.numeric)
cpeptide$cpeptide_AUC <- rowSums(cpeptide[,2:10])
cpeptide$cpeptide_mean <- cpeptide$cpeptide_AUC/9

glucagon <- gluc_wide %>% select(subject_id, contains("glucagon"),-starts_with("-10"))
glucagon <- glucagon[complete.cases(glucagon),]
glucagon[,2:10] <- sapply(glucagon[,2:10], as.numeric)
glucagon$glucagon_AUC <- rowSums(glucagon[,2:10])
glucagon$glucagon_mean <- glucagon$glucagon_AUC/9

gip <- gluc_wide %>% select(subject_id, contains("gip"),-starts_with("minus"))
gip <- gip[complete.cases(gip),]
gip[,2:10] <- sapply(gip[,2:10], as.numeric)
gip$gip_AUC <- rowSums(gip[,2:10])
gip$gip_mean <- gip$gip_AUC/9

glp1 <- gluc_wide %>% select(subject_id, contains("glp"),-starts_with("minus"))
glp1 <- glp1[complete.cases(glp1),]
glp1[,2:10] <- sapply(glp1[,2:10], as.numeric)
glp1$glp1_AUC <- rowSums(glp1[,2:10])
glp1$glp1_mean <- glp1$glp1_AUC/9

@@

auc_df$hypo_60 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])),"Yes","No")
auc_df$hypo_70 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])),"Yes","No")

auc_df$matsuda_ins <- 10000/sqrt(auc_df$`0_glucose`*auc_df$`0_insulin`*auc_df$glucose_mean*auc_df$insulin_mean)
auc_df$matsuda_cpep <- 20/(auc_df$`0_cpeptide`*auc_df$`0_glucose`)
auc_df$homa_ir <- auc_df$`0_glucose`*auc_df$`0_insulin`/405
auc_df$igi <- (auc_df$`30_insulin` - auc_df$`0_insulin`)/(auc_df$`30_glucose` - auc_df$`0_glucose`)
auc_df$igi[is.infinite(auc_df$igi)] <- NA
auc_df$cpep_index <- 100*auc_df$`0_cpeptide`/auc_df$`0_glucose`
auc_df$odi_ins <- auc_df$igi * auc_df$matsuda_ins
auc_df$odi_cpep <- auc_df$igi * auc_df$matsuda_cpep
auc_df$cpep_ins0 <- auc_df$`0_cpeptide`/auc_df$`0_insulin`
auc_df$cpep_auc_30 <- (auc_df$`0_cpeptide`+auc_df$`10_cpeptide`+auc_df$`20_cpeptide`+auc_df$`30_cpeptide`)/4
auc_df$ins_auc_30 <- (auc_df$`0_insulin`+auc_df$`10_insulin`+auc_df$`20_insulin`+auc_df$`30_insulin`)/4
auc_df$cpep_ins_auc_30 <- auc_df$cpep_auc_30/auc_df$ins_auc_30
auc_df$cpep_ins_auc <- auc_df$cpeptide_AUC/auc_df$insulin_AUC

@@

for (subject in auc_df$subject_id){
  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glucose")) 
  gluc_peak <- c(gluc_peak,max(sub_df))
  gluc_nadir <- c(gluc_nadir,min(sub_df))
  gluc_peak_time <- c(gluc_peak_time,
                      strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1])
  gluc_nadir_time <- c(gluc_nadir_time,
                       strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1])

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("insulin")) 
  ins_peak <- c(ins_peak,max(sub_df))
  ins_nadir <- c(ins_nadir,min(sub_df))
  ins_peak_time <- c(ins_peak_time,
                     strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1])
  ins_nadir_time <- c(ins_nadir_time,
                      strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1])

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("cpeptide")) 
  cpep_peak <- c(cpep_peak,max(sub_df))
  cpep_nadir <- c(cpep_nadir,min(sub_df))

  cppt_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  cpnt_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(cppt_temp) > 0){
    cpep_peak_time <- c(cpep_peak_time,
                        cppt_temp[[1]][1])
  }
  else {
    cpep_peak_time <- c(cpep_peak_time,
                        NA)
  }
  if(length(cpnt_temp) > 0){
    cpep_nadir_time <- c(cpep_nadir_time,
                         cpnt_temp[[1]][1])
  }
  else {
    cpep_nadir_time <- c(cpep_nadir_time,
                         NA)
  }

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glucagon")) 
  gluca_peak <- c(gluca_peak,max(sub_df))
  gluca_nadir <- c(gluca_nadir,min(sub_df))
  glucap_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glucan_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(glucap_temp) > 0){
    gluca_peak_time <- c(gluca_peak_time,
                         glucap_temp[[1]][1])
  }
  else {
    gluca_peak_time <- c(gluca_peak_time,
                         NA)
  }
  if(length(glucan_temp) > 0){
    gluca_nadir_time <- c(gluca_nadir_time,
                          glucan_temp[[1]][1])
  }
  else {
    gluca_nadir_time <- c(gluca_nadir_time,
                          NA)
  }

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("gip")) 
  gip_peak <- c(gip_peak,max(sub_df))
  gip_nadir <- c(gip_nadir,min(sub_df))
  gipp_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  gipn_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(gipp_temp) > 0){
    gip_peak_time <- c(gip_peak_time,
                       gipp_temp[[1]][1])
  }
  else {
    gip_peak_time <- c(gip_peak_time,
                       NA)
  }
  if(length(gipn_temp) > 0){
    gip_nadir_time <- c(gip_nadir_time,
                        gipn_temp[[1]][1])
  }
  else {
    gip_nadir_time <- c(gip_nadir_time,
                        NA)
  }

  sub_df <- auc_df %>% filter(subject_id == subject) %>% select(ends_with("glp1")) 
  glp1_peak <- c(glp1_peak,max(sub_df))
  glp1_nadir <- c(glp1_nadir,min(sub_df))
  glp1p_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glp1n_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if(length(gipp_temp) > 0){
    glp1_peak_time <- c(glp1_peak_time,
                        glp1p_temp[[1]][1])
  }
  else {
    glp1_peak_time <- c(glp1_peak_time,
                        NA)
  }
  if(length(glp1n_temp) > 0){
    glp1_nadir_time <- c(glp1_nadir_time,
                         glp1n_temp[[1]][1])
  }
  else {
    glp1_nadir_time <- c(glp1_nadir_time,
                         NA)
  }
}

@@

table_dat <- merge(new_gluc,gluc_aucs)
table_dat <- merge(table_dat, auc_df %>% select(subject_id, 
                                                matsuda_ins, 
                                                matsuda_cpep, 
                                                homa_ir, 
                                                igi, 
                                                cpep_index,
                                                odi_ins,
                                                odi_cpep,
                                                cpep_ins0,
                                                cpep_ins_auc_30,
                                                cpep_ins_auc,
                                                `0_glucose`,
                                                `0_insulin`,
                                                `0_cpeptide`,
                                                `0_glucagon`,
                                                `0_gip`,
                                                `0_glp1`,
                                                gluc_peak,
                                                gluc_nadir,
                                                gluc_peak_time,
                                                gluc_nadir_time,
                                                ins_peak,
                                                ins_nadir,
                                                ins_peak_time,
                                                ins_nadir_time,
                                                cpep_peak,
                                                cpep_nadir,
                                                cpep_peak_time,
                                                cpep_nadir_time,
                                                gluca_peak,
                                                gluca_nadir,
                                                gluca_peak_time,
                                                gluca_nadir_time,
                                                gip_peak,
                                                gip_nadir,
                                                gip_peak_time,
                                                gip_nadir_time,
                                                glp1_peak,
                                                glp1_nadir,
                                                glp1_peak_time,
                                                glp1_nadir_time
), all = T)

cgm <- dat %>% select(subject_id,
                      date_of_visit,
                      num_days_good_data:interquartile_range
) %>% 
  filter(subject_id %in% keep_ids) %>% 
  filter(!is.na(num_days_good_data)) %>%
  unique()

cgm$daytime_avg_sensor_glucose <- round(as.numeric(cgm$daytime_avg_sensor_glucose),4)
cgm$percent_time_over_140 <- round(as.numeric(cgm$percent_time_over_140),4)
cgm$percent_time_over_180 <- round(as.numeric(cgm$percent_time_over_180),4)
cgm$daytime_sd <- round(as.numeric(cgm$daytime_sd),4)
cgm$cov <- round(as.numeric(cgm$cov),4)
cgm$mag <- round(as.numeric(cgm$mag),4)
cgm <- unique(cgm)
cgm <- cgm[!duplicated(cgm[,1:2]),]
cgm <- cgm[!duplicated(cgm$subject_id),]

@@

dat$mutation_severity[dat$cf_mutation_1_class <=3 | dat$cf_mutation_2_class <=3] <-  "Severe"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 4 | 
                                                        dat$cf_mutation_2_class == 4 |
                                                        dat$cf_mutation_1_class <= 5 | 
                                                        dat$cf_mutation_2_class <=5)] <-  "Mild"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 6 | 
                                                        dat$cf_mutation_2_class == 6)] <-  "Other"

dat_demo <- dat %>% select(subject_id, 
                           date_of_visit,
                           group_category,
                           gender, 
                           race_ethnicity, 
                           tanner_stage,
                           weight, 
                           height, 
                           bmi, 
                           bmi_z_score, 
                           mutation_severity, 
                           fev1_percent,
                           fvc,
                           poc_a1c,
                           cf_pancreatic,
                           cf_insulin_use,
                           g_tube_feeds)
dat_demo <- dat_demo %>% filter(subject_id %in% keep_ids)

new_demo <- dat_demo[0,]
for (id in unique(dat_demo$subject_id)){
  id_df <- dat_demo %>% filter(subject_id == id) %>% arrange(date_of_visit) %>% na.locf(na.rm = FALSE)
  id_df <- id_df %>% arrange(desc(date_of_visit)) %>% na.locf(na.rm = FALSE)
  new_demo <- rbind(new_demo,id_df)
}

@@

dat_demo <- dat_demo[!duplicated(dat_demo$subject_id),]

>>>>

---
title: "Hypoglycemia Analysis - CFRD REMOVED"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    fig_height: 8
  pdf_document:
    toc: yes
---

@@

```{r}
getAUCs <- function(data, subject, time, vars) {
  all_return <- NULL

  for (var in vars) {
    subs <- unique(data[[subject]])
    return <- NULL
    for (sub in subs) {
      try({
        df <- data[data[[subject]] == sub, ]
        ogtt_0 <- df[df[[time]] == 0, var]
        ogtt_30 <- df[df[[time]] == 30, var]
        ogtt_60 <- df[df[[time]] == 60, var]
        ogtt_90 <- df[df[[time]] == 90, var]
        ogtt_120 <- df[df[[time]] == 120, var]
        ogtt_auc <- (ogtt_30[[var]] + ogtt_60[[var]] + ogtt_90[[var]]) / 
          2 + (ogtt_0[[var]] + ogtt_120[[var]]) / 4
        if (length(ogtt_auc) > 0) {
          row <- c(sub, ogtt_auc)
          return <- rbind(return, row)
        }
      })
    }
    return <- as.data.frame(return)
    names(return) <- c(subject, paste0(var, "_auc"))
    if (is.null(all_return)) {
      all_return <- return
    } else {
      all_return <- merge(all_return, return, all = T)
    }
  }
  all_return
}
```

```{r, warning = F, message = F}
dat <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan/Hypoglycemia/Background/Previous Analyses/FsOGTT/data/CGMInHealthyControls_DATA_2021-05-25_0918.csv", na = "-999")

dat$group_category[dat$group_category == "1"] <- "Healthy"
dat$group_category[dat$group_category == "2"] <- "CF-NGT"
dat$group_category[dat$group_category == "3"] <- "CF-AG"
dat$group_category[dat$group_category == "4"] <- "CFRD"
dat$group_category <- factor(dat$group_category, levels = c("Healthy", "CF-NGT", "CF-AG", "CFRD"))

@@

dat <- dat[dat$date_of_visit != "", ]

@@

gluc <- dat %>% select(
  subject_id,
  date_of_visit,
  age,
  gender,
  race_ethnicity,
  starts_with("min"),
  -starts_with("min_spent"),
  -contains("sensor"),
  -contains("active"),
  -contains("glp"),
  -contains("gip")
)
gluc <- merge(gluc, dat %>% select(subject_id, date_of_visit, age, starts_with("ucd_")))
gluc <- unique(gluc)
gluc <- gluc[!is.na(gluc$min_30_glucose), ]
gluc[] <- lapply(gluc, gsub, pattern = "<", replacement = "")

gluc_names <- names(gluc)
gluc_names <- gsub("min_", "", gluc_names)
gluc_names <- gsub("minute_", "", gluc_names)
gluc_names <- gsub("minus_", "-", gluc_names)
gluc_names <- gsub("ucd_", "", gluc_names)
gluc_names <- gsub("_total", "", gluc_names)
gluc_names <- gsub("glp_1", "_glp1", gluc_names)
gluc_names <- gsub("c_peptide", "cpeptide", gluc_names)
names(gluc) <- gluc_names
gluc[gluc == -999.99] <- NA
gluc <- unique(gluc)
new_gluc <- gluc[0, ]

for (sub in unique(gluc$subject_id)) {
  df <- gluc[gluc$subject_id == sub, ]
  if (nrow(df) > 1) {
    df <- df %>% arrange(`0_gip`, desc(date_of_visit))
    df <- df[1, ]
  }
  new_gluc <- rbind(new_gluc, df)
}

gluc_wide <- new_gluc %>% select(-date_of_visit)
gluc <- gluc_wide %>% pivot_longer(
  cols = -c(subject_id),
  names_to = c("time", ".value"), names_sep = "_"
)

# gluc <- gluc[!is.na(gluc$glucose) | !is.na(gluc$glucagon) | !is.na(gluc$insulin)| !is.na(gluc$cpeptide),]
# gluc <- gluc[!is.na(gluc$glucagon),]
gluc <- gluc[gluc$time != "-10", ]
gluc$time <- as.factor(as.numeric(gluc$time))

@@

aucs <- getAUCs(gluc, "subject_id", "time", c("glucose", "glucagon", "insulin", "cpeptide", "gip", "glp1"))

df_aucs <- merge(aucs, unique(gluc %>% select(subject_id)))
df_aucs$glucagon_auc <- as.numeric(df_aucs$glucagon_auc)
df_aucs$glucose_auc <- as.numeric(df_aucs$glucose_auc)
df_aucs$insulin_auc <- as.numeric(df_aucs$insulin_auc)
df_aucs$cpeptide_auc <- as.numeric(df_aucs$cpeptide_auc)
df_aucs$gip_auc <- as.numeric(df_aucs$gip_auc)
df_aucs$glp1_auc <- as.numeric(df_aucs$glp1_auc)

df_aucs <- df_aucs[!is.na(df_aucs$glucagon_auc) | !is.na(df_aucs$glucose_auc) | !is.na(df_aucs$insulin_auc) | !is.na(df_aucs$cpeptide_auc), ]

@@

gluc$hypo_60 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])), "Yes", "No")
gluc$hypo_70 <- ifelse(gluc$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])), "Yes", "No")

gluc_aucs <- merge(gluc_aucs, unique(gluc %>% select(subject_id, hypo_60, hypo_70)))

glucose <- gluc_wide %>% select(subject_id, contains("glucose"), -starts_with("-10"))
glucose <- glucose[complete.cases(glucose), ]
glucose[, 2:10] <- sapply(glucose[, 2:10], as.numeric)
glucose$glucose_AUC <- rowSums(glucose[, 2:10])
glucose$glucose_mean <- glucose$glucose_AUC / 9

insulin <- gluc_wide %>% select(subject_id, contains("insulin"), -starts_with("-10"))
insulin <- insulin[complete.cases(insulin), ]
insulin[, 2:10] <- sapply(insulin[, 2:10], as.numeric)
insulin$insulin_AUC <- rowSums(insulin[, 2:10])
insulin$insulin_mean <- insulin$insulin_AUC / 9

cpeptide <- gluc_wide %>% select(subject_id, contains("cpeptide"), -starts_with("-10"))
cpeptide <- cpeptide[complete.cases(cpeptide), ]
cpeptide[, 2:10] <- sapply(cpeptide[, 2:10], as.numeric)
cpeptide$cpeptide_AUC <- rowSums(cpeptide[, 2:10])
cpeptide$cpeptide_mean <- cpeptide$cpeptide_AUC / 9

glucagon <- gluc_wide %>% select(subject_id, contains("glucagon"), -starts_with("-10"))
glucagon <- glucagon[complete.cases(glucagon), ]
glucagon[, 2:10] <- sapply(glucagon[, 2:10], as.numeric)
glucagon$glucagon_AUC <- rowSums(glucagon[, 2:10])
glucagon$glucagon_mean <- glucagon$glucagon_AUC / 9

gip <- gluc_wide %>% select(subject_id, contains("gip"), -starts_with("minus"))
gip <- gip[complete.cases(gip), ]
gip[, 2:10] <- sapply(gip[, 2:10], as.numeric)
gip$gip_AUC <- rowSums(gip[, 2:10])
gip$gip_mean <- gip$gip_AUC / 9

glp1 <- gluc_wide %>% select(subject_id, contains("glp"), -starts_with("minus"))
glp1 <- glp1[complete.cases(glp1), ]
glp1[, 2:10] <- sapply(glp1[, 2:10], as.numeric)
glp1$glp1_AUC <- rowSums(glp1[, 2:10])
glp1$glp1_mean <- glp1$glp1_AUC / 9

@@

auc_df$hypo_60 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 60])), "Yes", "No")
auc_df$hypo_70 <- ifelse(auc_df$subject_id %in% na.omit(unique(gluc$subject_id[gluc$glucose < 70])), "Yes", "No")

auc_df$matsuda_ins <- 10000 / sqrt(auc_df$`0_glucose` * auc_df$`0_insulin` * auc_df$glucose_mean * auc_df$insulin_mean)
auc_df$matsuda_cpep <- 20 / (auc_df$`0_cpeptide` * auc_df$`0_glucose`)
auc_df$homa_ir <- auc_df$`0_glucose` * auc_df$`0_insulin` / 405
auc_df$igi <- (auc_df$`30_insulin` - auc_df$`0_insulin`) / (auc_df$`30_glucose` - auc_df$`0_glucose`)
auc_df$igi[is.infinite(auc_df$igi)] <- NA
auc_df$cpep_index <- 100 * auc_df$`0_cpeptide` / auc_df$`0_glucose`
auc_df$odi_ins <- auc_df$igi * auc_df$matsuda_ins
auc_df$odi_cpep <- auc_df$igi * auc_df$matsuda_cpep
auc_df$cpep_ins0 <- auc_df$`0_cpeptide` / auc_df$`0_insulin`
auc_df$cpep_auc_30 <- (auc_df$`0_cpeptide` + auc_df$`10_cpeptide` + auc_df$`20_cpeptide` + auc_df$`30_cpeptide`) / 4
auc_df$ins_auc_30 <- (auc_df$`0_insulin` + auc_df$`10_insulin` + auc_df$`20_insulin` + auc_df$`30_insulin`) / 4
auc_df$cpep_ins_auc_30 <- auc_df$cpep_auc_30 / auc_df$ins_auc_30
auc_df$cpep_ins_auc <- auc_df$cpeptide_AUC / auc_df$insulin_AUC

@@

for (subject in auc_df$subject_id) {
  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("glucose"))
  gluc_peak <- c(gluc_peak, max(sub_df))
  gluc_nadir <- c(gluc_nadir, min(sub_df))
  gluc_peak_time <- c(
    gluc_peak_time,
    strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1]
  )
  gluc_nadir_time <- c(
    gluc_nadir_time,
    strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1]
  )

  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("insulin"))
  ins_peak <- c(ins_peak, max(sub_df))
  ins_nadir <- c(ins_nadir, min(sub_df))
  ins_peak_time <- c(
    ins_peak_time,
    strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")[[1]][1]
  )
  ins_nadir_time <- c(
    ins_nadir_time,
    strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")[[1]][1]
  )

  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("cpeptide"))
  cpep_peak <- c(cpep_peak, max(sub_df))
  cpep_nadir <- c(cpep_nadir, min(sub_df))

  cppt_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  cpnt_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(cppt_temp) > 0) {
    cpep_peak_time <- c(
      cpep_peak_time,
      cppt_temp[[1]][1]
    )
  } else {
    cpep_peak_time <- c(
      cpep_peak_time,
      NA
    )
  }
  if (length(cpnt_temp) > 0) {
    cpep_nadir_time <- c(
      cpep_nadir_time,
      cpnt_temp[[1]][1]
    )
  } else {
    cpep_nadir_time <- c(
      cpep_nadir_time,
      NA
    )
  }

  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("glucagon"))
  gluca_peak <- c(gluca_peak, max(sub_df))
  gluca_nadir <- c(gluca_nadir, min(sub_df))
  glucap_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glucan_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(glucap_temp) > 0) {
    gluca_peak_time <- c(
      gluca_peak_time,
      glucap_temp[[1]][1]
    )
  } else {
    gluca_peak_time <- c(
      gluca_peak_time,
      NA
    )
  }
  if (length(glucan_temp) > 0) {
    gluca_nadir_time <- c(
      gluca_nadir_time,
      glucan_temp[[1]][1]
    )
  } else {
    gluca_nadir_time <- c(
      gluca_nadir_time,
      NA
    )
  }

  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("gip"))
  gip_peak <- c(gip_peak, max(sub_df))
  gip_nadir <- c(gip_nadir, min(sub_df))
  gipp_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  gipn_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(gipp_temp) > 0) {
    gip_peak_time <- c(
      gip_peak_time,
      gipp_temp[[1]][1]
    )
  } else {
    gip_peak_time <- c(
      gip_peak_time,
      NA
    )
  }
  if (length(gipn_temp) > 0) {
    gip_nadir_time <- c(
      gip_nadir_time,
      gipn_temp[[1]][1]
    )
  } else {
    gip_nadir_time <- c(
      gip_nadir_time,
      NA
    )
  }

  sub_df <- auc_df %>%
    filter(subject_id == subject) %>%
    select(ends_with("glp1"))
  glp1_peak <- c(glp1_peak, max(sub_df))
  glp1_nadir <- c(glp1_nadir, min(sub_df))
  glp1p_temp <- strsplit(names(sub_df)[which(sub_df == max(sub_df))], "_")
  glp1n_temp <- strsplit(names(sub_df)[which(sub_df == min(sub_df))], "_")
  if (length(gipp_temp) > 0) {
    glp1_peak_time <- c(
      glp1_peak_time,
      glp1p_temp[[1]][1]
    )
  } else {
    glp1_peak_time <- c(
      glp1_peak_time,
      NA
    )
  }
  if (length(glp1n_temp) > 0) {
    glp1_nadir_time <- c(
      glp1_nadir_time,
      glp1n_temp[[1]][1]
    )
  } else {
    glp1_nadir_time <- c(
      glp1_nadir_time,
      NA
    )
  }
}

@@

table_dat <- merge(new_gluc, gluc_aucs)
table_dat <- merge(table_dat, auc_df %>% select(
  subject_id,
  matsuda_ins,
  matsuda_cpep,
  homa_ir,
  igi,
  cpep_index,
  odi_ins,
  odi_cpep,
  cpep_ins0,
  cpep_ins_auc_30,
  cpep_ins_auc,
  `0_glucose`,
  `0_insulin`,
  `0_cpeptide`,
  `0_glucagon`,
  `0_gip`,
  `0_glp1`,
  gluc_peak,
  gluc_nadir,
  gluc_peak_time,
  gluc_nadir_time,
  ins_peak,
  ins_nadir,
  ins_peak_time,
  ins_nadir_time,
  cpep_peak,
  cpep_nadir,
  cpep_peak_time,
  cpep_nadir_time,
  gluca_peak,
  gluca_nadir,
  gluca_peak_time,
  gluca_nadir_time,
  gip_peak,
  gip_nadir,
  gip_peak_time,
  gip_nadir_time,
  glp1_peak,
  glp1_nadir,
  glp1_peak_time,
  glp1_nadir_time
), all = T)

cgm <- dat %>%
  select(
    subject_id,
    date_of_visit,
    num_days_good_data:interquartile_range
  ) %>%
  filter(subject_id %in% keep_ids) %>%
  filter(!is.na(num_days_good_data)) %>%
  unique()

cgm$daytime_avg_sensor_glucose <- round(as.numeric(cgm$daytime_avg_sensor_glucose), 4)
cgm$percent_time_over_140 <- round(as.numeric(cgm$percent_time_over_140), 4)
cgm$percent_time_over_180 <- round(as.numeric(cgm$percent_time_over_180), 4)
cgm$daytime_sd <- round(as.numeric(cgm$daytime_sd), 4)
cgm$cov <- round(as.numeric(cgm$cov), 4)
cgm$mag <- round(as.numeric(cgm$mag), 4)
cgm <- unique(cgm)
cgm <- cgm[!duplicated(cgm[, 1:2]), ]
cgm <- cgm[!duplicated(cgm$subject_id), ]

@@

dat$mutation_severity[dat$cf_mutation_1_class <= 3 | dat$cf_mutation_2_class <= 3] <- "Severe"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 4 |
  dat$cf_mutation_2_class == 4 |
  dat$cf_mutation_1_class <= 5 |
  dat$cf_mutation_2_class <= 5)] <- "Mild"
dat$mutation_severity[is.na(dat$mutation_severity) & (dat$cf_mutation_1_class == 6 |
  dat$cf_mutation_2_class == 6)] <- "Other"

dat_demo <- dat %>% select(
  subject_id,
  date_of_visit,
  group_category,
  gender,
  race_ethnicity,
  tanner_stage,
  weight,
  height,
  bmi,
  bmi_z_score,
  mutation_severity,
  fev1_percent,
  fvc,
  poc_a1c,
  cf_pancreatic,
  cf_insulin_use,
  g_tube_feeds
)
dat_demo <- dat_demo %>% filter(subject_id %in% keep_ids)

new_demo <- dat_demo[0, ]
for (id in unique(dat_demo$subject_id)) {
  id_df <- dat_demo %>%
    filter(subject_id == id) %>%
    arrange(date_of_visit) %>%
    na.locf(na.rm = FALSE)
  id_df <- id_df %>%
    arrange(desc(date_of_visit)) %>%
    na.locf(na.rm = FALSE)
  new_demo <- rbind(new_demo, id_df)
}

@@

dat_demo <- dat_demo[!duplicated(dat_demo$subject_id), ]

table_dat <- merge(table_dat, dat_demo, all = T)
